\thesistitle

### 2019-08-11 sections/conclusion.tex
Proper capitalization for some concepts

Refactoring complex methods can largely be done automatically. Also, many of the results of this study can be reused. To assess an automated refactoring opportunity for complex methods, we should assess which parts of methods can be split to end up with parts of similar complexity. For this, our research can be used to assess whether a given piece of code can be extracted to a new method (see section \ref{sec:refactorability}).

@@

\subsection{Type 4 clones}
In section \ref{chap:backgroundclonetypes} we introduced the clone types that are used in literature. Of these clone types, we considered type 1, 2 and 3 for refactoring. We chose not to consider type 4 clones because they appear far less in source code and are more difficult to detect and refactor. However, that does not mean that these clones should not be refactored.

>>>>

Refactoring complex methods can largely be done automatically. Also, many of the results of this study can be reused. To assess an automated refactoring opportunity for complex methods, we should assess which parts of methods can be split to end up with parts of similar complexity. For this, our research can be used to assess whether a given piece of code can be extracted to a new method (see Section~\ref{sec:refactorability}).

@@

\subsection{Type 4 clones}
In Section~\ref{chap:backgroundclonetypes} we introduced the clone types that are used in literature. Of these clone types, we considered type 1, 2 and 3 for refactoring. We chose not to consider type 4 clones because they appear far less in source code and are more difficult to detect and refactor. However, that does not mean that these clones should not be refactored.

### 2019-08-11 sections/discussion.tex
Proper capitalization for some concepts

\section{Clone Type Definitions}
In this study, we proposed a set of clone type definitions for which a refactoring opportunity is known. We based these clone type definitions on the clone type definitions that are commonly used in literature. The design of these clone type definitions entails some decisions that have a large impact on the clones we considered for refactoring in this study. In this section we discuss our clone type definitions as proposed in section \ref{sec:rtypes}.

@@

One method we think can be used to find more refactoring opportunities is to allow variability in expressions that have/return the same type. If expressions have/return the same type, they can be extracted to a parameter and the corresponding expression can be passed as a parameter. An example of this is displayed in figure \ref{fig:samereturn}. The only thing to watch out for is method that have side effects. Because methods may be executed in another point during execution, this might affect the functionality of the code.

@@

\section{Clone Context Analysis}
In section \ref{chap:contextsetup} we introduced the categories we defined for mapping the context of clones. In this section, we discuss this together with the related experiments.

\subsection{Refactorability} \label{sec:discussrefactorability}
In section \ref{sec:refactorability} we introduced to what extent clones can be refactored through method extraction. Because we strived to get results fast, we excluded categories that could not not be directly refactored through method extraction. However, with a few transformations or further considerations it might be possible to make these clones refactorable. In this section we will highlight a few of these categories which we believe to be refactorable through method extraction with a bit more effort.

>>>>

\section{Clone Type Definitions}
In this study, we proposed a set of clone type definitions for which a refactoring opportunity is known. We based these clone type definitions on the clone type definitions that are commonly used in literature. The design of these clone type definitions entails some decisions that have a large impact on the clones we considered for refactoring in this study. In this section we discuss our clone type definitions as proposed in Section~\ref{sec:rtypes}.

@@

One method we think can be used to find more refactoring opportunities is to allow variability in expressions that have/return the same type. If expressions have/return the same type, they can be extracted to a parameter and the corresponding expression can be passed as a parameter. An example of this is displayed in Figure~\ref{fig:samereturn}. The only thing to watch out for is method that have side effects. Because methods may be executed in another point during execution, this might affect the functionality of the code.

@@

\section{Clone Context Analysis}
In Section~\ref{chap:contextsetup} we introduced the categories we defined for mapping the context of clones. In this section, we discuss this together with the related experiments.

\subsection{Refactorability} \label{sec:discussrefactorability}
In Section~\ref{sec:refactorability} we introduced to what extent clones can be refactored through method extraction. Because we strived to get results fast, we excluded categories that could not not be directly refactored through method extraction. However, with a few transformations or further considerations it might be possible to make these clones refactorable. In this section we will highlight a few of these categories which we believe to be refactorable through method extraction with a bit more effort.

### 2019-08-11 sections/research.tex
Ana
Proper capitalization for some concepts

\chapter{Defining refactoring-oriented clone types}\label{chap:clonetypes}
In section \ref{chap:backgroundclonetypes} we introduced the four clone types as defined in literature. These simple definitions are suitable for analysis of a codebase. Their detection results in simple to understand numbers to argue about a codebase. However, these clone types have a few flaws which makes it hard to argue to what extend two fragments of code can and should be refactored. For each of type 1-3 clones~\cite{roy2007survey} we list our solutions to their shortcomings to increase the chance that we can refactor the clone while improving the design.

@@

\section{Shortcomings of clone types}
Clone types 1-3 (further explained in section \ref{chap:backgroundclonetypes}) allow reasoning about the duplication in a software system. Clones by these definitions can relatively easily and efficiently be detected. This has allowed for large scale analyses of duplication~\cite{livieri2007very}. However, these clone type definitions have shortcomings which makes the clones detected in correspondence with these definitions less valuable for (automated) refactoring purposes.

@@

In the example in figure \ref{fig:type1}, we see a type 1 clone displaying two clone classes. Defining an automatic way to refactor these clone classes is nearly impossible, as both cloned fragments describe different functional concepts. The first cloned fragment is a method that adds something to a \texttt{List}. However, the \texttt{List} objects to which something is added are different. Looking at the \texttt{import} statement above the class, one fragment uses the \texttt{java.util.List} and the other uses the \texttt{java.awt.List}. Both happen to have an \texttt{add} method, but apart from that their implementation is completely different.

@@

The example in figure \ref{fig:type2} shows a type 2 clone class. Both methods are, except for their matching structure, completely different in functionality. They operate on different types, call different methods, return different things, etc. Having such a method flagged as a clone does not provide much useful information.

@@

\subsection{Type 1R clones} \label{sec:type1r}
To solve the issues identified in section \ref{sec:type1}, we introduce an alternative definition: cloned fragments have to be both textually \textit{and} functionally equal. Like type 1 clones, type 1R clones do not consider comments and whitespace. Therefore, T1R clones are a subset of type 1 clones.

@@

\subsubsection{Variables}
In typed programming languages, each variable declaration should declare a name and a type. When we reference a variable, we only use its name. If, in different code fragments, we use variables with the same name but different types, the code can be functionally unequal but still textually equal. As an example, see the code in figure \ref{fig:type2variables}.

@@

The body of both methods in figure \ref{fig:type2variables} is equal. However, their functionality is not. The first method adds two numbers together and the other concatenates an integer to a String.

@@

\subsection{Type 2R clones} \label{sec:type2r}
Type 2R clones are modelled after type 2 clones, which allow any change in identifiers, literals, types, layout, and comments. For refactoring purposes, this definition is unsuitable; if we allow any change in identifiers, literals, and types, we cannot distinguish between different variables, different types and different method calls anymore. This could render two methods that have an entirely different functionality as clones (as shown in figure \ref{fig:type2} previously).

@@

Allowing variability between expressions includes a tradeoff. When many expressions vary between cloned the refactored method might require many additional arguments. Because of that, we constrain this variability by a threshold, which is explained in more detail in section \ref{sec:variabilitythreshold}.%to limit the negative effect on system design of variability in these expressions after refactoring.

@@

The example in figure \ref{fig:type2variablerefactoringwithtradeoff} shows the same example as figure \ref{fig:type2variablerefactoringnotradeoff}, however this time there is a tradeoff. The same variables are used, but they are not used in the same places in cloned fragments. In one fragment we use only one variable, in the other we use both. Because of that, we require an extra argument. Because of this, this variability falls under the threshold described in section \label{sec:variabilitythreshold}, whereas the example in figure \ref{fig:type2variablerefactoringnotradeoff} does not.

@@

Consider the example in figure \ref{fig:type2literal}. In this example, the two methods have two literals that differ between them. We can perform an ``Extract method'' refactoring on these to get the result that is displayed on the right. In this process, we create a method parameter for the corresponding literal.

@@

Type 2R clones allow called methods to vary as long as they have the same argument types and return type. As with type 1R clones, these types are compared using their fully qualified identifiers. An example of this is shown in figure \ref{fig:type2method}. In this example, we have two methods (\texttt{System.out.println} and \texttt{myFancyPrint}). We use the ``Extract Method'' refactoring method to extract a new method and use a parameter to pass the used method.

The method call variability property of type 2R clones imply that type 2R clones are not a subset of type 2 clones. Because methods calls can have a different structure, type 2R clones can be structurally slightly different. The example as shown in figure \ref{fig:type2method} can be a type 2R clone (dependent on the thresholds used), but is not a type 2 clone.

@@

\subsection{Type 3R clones}\label{sec:type3r}
Type 3 clones allow any change in statements, often bounded by a similarity threshold. This means that type 3 clones allow the inclusion of a statement that is not detected by type 1 or 2 clone detection. When looking at how we can refactor a statement that is not included by one clone instance but is in another, we find that we require a conditional block to make up for the difference in statements. See figure \ref{fig:type3} for an example of such a clone.

@@

In figure \ref{fig:type3} a single statement is added. This statement is found in between cloned lines. There is a \textit{gap} of non-cloned lines in between two clone classes. The following rules apply to this gap:

\begin{itemize}
  \item \textbf{The difference in statements must bridge a gap between two valid clones}. This entails that, different from type 3 clones, the difference in statements cannot be at the beginning or the end of a cloned block. It is rather somewhere within, as it must bridge two existent clones.
  \item \textbf{The size of the gap between two clones is limited by a threshold.} This threshold is calculated by taking the percentage of the number of statements in the gap over the number of statements that both clones that are being bridged span.
  \item \textbf{The gap may not span a partial block.} To make sure that the T3R clone can be refactored, we do not allow the gap to span a part of a block. This is further explained in section \ref{sec:t3rpartialblock}.
\end{itemize}

@@

As an example, consider the code fragment in figure \ref{fig:gappedclones}.

@@

When applying this formula with the example given in figure \ref{fig:gappedclones}, we get a gap size of 50\%:

@@

\subsubsection{The gap may not span a partial block} \label{sec:t3rpartialblock}
Apart from this threshold, the gap in between clones may not span over different blocks. Look at figure \ref{fig:type3invalid} for an example of this. We cannot refactor both statements into a single conditional block. We could however use two conditional blocks, but due to the detrimental effect on the design of the code (as each conditional block adds a certain complexity), we decided not to allow this for type 3R clones.

@@

\begin{equation}\label{eq:typerelation}
\text{Cloned nodes} = T1R \subseteq T2R \subseteq T3R
\end{equation}

>>>>

\chapter{Defining refactoring-oriented clone types}\label{chap:clonetypes}
In Section~\ref{chap:backgroundclonetypes} we introduced the four clone types as defined in literature. These simple definitions are suitable for analysis of a codebase. Their detection results in simple to understand numbers to argue about a codebase. However, these clone types have a few flaws which makes it hard to argue to what extend two fragments of code can and should be refactored. For each of type 1-3 clones~\cite{roy2007survey} we list our solutions to their shortcomings to increase the chance that we can refactor the clone while improving the design.

@@

\section{Shortcomings of clone types}
Clone types 1-3 (further explained in Section~\ref{chap:backgroundclonetypes}) allow reasoning about the duplication in a software system. Clones by these definitions can relatively easily and efficiently be detected. This has allowed for large scale analyses of duplication~\cite{livieri2007very}. However, these clone type definitions have shortcomings which makes the clones detected in correspondence with these definitions less valuable for (automated) refactoring purposes.

@@

In the example in Figure~\ref{fig:type1}, we see a type 1 clone displaying two clone classes. Defining an automatic way to refactor these clone classes is nearly impossible, as both cloned fragments describe different functional concepts. The first cloned fragment is a method that adds something to a \texttt{List}. However, the \texttt{List} objects to which something is added are different. Looking at the \texttt{import} statement above the class, one fragment uses the \texttt{java.util.List} and the other uses the \texttt{java.awt.List}. Both happen to have an \texttt{add} method, but apart from that their implementation is completely different.

@@

The example in Figure~\ref{fig:type2} shows a type 2 clone class. Both methods are, except for their matching structure, completely different in functionality. They operate on different types, call different methods, return different things, etc. Having such a method flagged as a clone does not provide much useful information.

@@

\subsection{Type 1R clones} \label{sec:type1r}
To solve the issues identified in Section~\ref{sec:type1}, we introduce an alternative definition: cloned fragments have to be both textually \textit{and} functionally equal. Like type 1 clones, type 1R clones do not consider comments and whitespace. Therefore, T1R clones are a subset of type 1 clones.

@@

\subsubsection{Variables}
In typed programming languages, each variable declaration should declare a name and a type. When we reference a variable, we only use its name. If, in different code fragments, we use variables with the same name but different types, the code can be functionally unequal but still textually equal. As an example, see the code in Figure~\ref{fig:type2variables}.

@@

The body of both methods in Figure~\ref{fig:type2variables} is equal. However, their functionality is not. The first method adds two numbers together and the other concatenates an integer to a String.

@@

\subsection{Type 2R clones} \label{sec:type2r}
Type 2R clones are modelled after type 2 clones, which allow any change in identifiers, literals, types, layout, and comments. For refactoring purposes, this definition is unsuitable; if we allow any change in identifiers, literals, and types, we cannot distinguish between different variables, different types and different method calls anymore. This could render two methods that have an entirely different functionality as clones (as shown in Figure~\ref{fig:type2} previously).

@@

Allowing variability between expressions includes a tradeoff. When many expressions vary between cloned the refactored method might require many additional arguments. Because of that, we constrain this variability by a threshold, which is explained in more detail in Section~\ref{sec:variabilitythreshold}.%to limit the negative effect on system design of variability in these expressions after refactoring.

@@

The example in Figure~\ref{fig:type2variablerefactoringwithtradeoff} shows the same example as Figure~\ref{fig:type2variablerefactoringnotradeoff}, however this time there is a tradeoff. The same variables are used, but they are not used in the same places in cloned fragments. In one fragment we use only one variable, in the other we use both. Because of that, we require an extra argument. Because of this, this variability falls under the threshold described in section \label{sec:variabilitythreshold}, whereas the example in Figure~\ref{fig:type2variablerefactoringnotradeoff} does not.

@@

Consider the example in Figure~\ref{fig:type2literal}. In this example, the two methods have two literals that differ between them. We can perform an ``Extract method'' refactoring on these to get the result that is displayed on the right. In this process, we create a method parameter for the corresponding literal.

@@

Type 2R clones allow called methods to vary as long as they have the same argument types and return type. As with type 1R clones, these types are compared using their fully qualified identifiers. An example of this is shown in Figure~\ref{fig:type2method}. In this example, we have two methods (\texttt{System.out.println} and \texttt{myFancyPrint}). We use the ``Extract Method'' refactoring method to extract a new method and use a parameter to pass the used method.

The method call variability property of type 2R clones imply that type 2R clones are not a subset of type 2 clones. Because methods calls can have a different structure, type 2R clones can be structurally slightly different. The example as shown in Figure~\ref{fig:type2method} can be a type 2R clone (dependent on the thresholds used), but is not a type 2 clone.

@@

\subsection{Type 3R clones}\label{sec:type3r}
Type 3 clones allow any change in statements, often bounded by a similarity threshold. This means that type 3 clones allow the inclusion of a statement that is not detected by type 1 or 2 clone detection. When looking at how we can refactor a statement that is not included by one clone instance but is in another, we find that we require a conditional block to make up for the difference in statements. See Figure~\ref{fig:type3} for an example of such a clone.

@@

In Figure~\ref{fig:type3} a single statement is added. This statement is found in between cloned lines. There is a \textit{gap} of non-cloned lines in between two clone classes. The following rules apply to this gap:

\begin{itemize}
  \item \textbf{The difference in statements must bridge a gap between two valid clones}. This entails that, different from type 3 clones, the difference in statements cannot be at the beginning or the end of a cloned block. It is rather somewhere within, as it must bridge two existent clones.
  \item \textbf{The size of the gap between two clones is limited by a threshold.} This threshold is calculated by taking the percentage of the number of statements in the gap over the number of statements that both clones that are being bridged span.
  \item \textbf{The gap may not span a partial block.} To make sure that the T3R clone can be refactored, we do not allow the gap to span a part of a block. This is further explained in Section~\ref{sec:t3rpartialblock}.
\end{itemize}

@@

As an example, consider the code fragment in Figure~\ref{fig:gappedclones}.

@@

When applying this formula with the example given in Figure~\ref{fig:gappedclones}, we get a gap size of 50\%:

@@

\subsubsection{The gap may not span a partial block} \label{sec:t3rpartialblock}
Apart from this threshold, the gap in between clones may not span over different blocks. Look at Figure~\ref{fig:type3invalid} for an example of this. We cannot refactor both statements into a single conditional block. We could however use two conditional blocks, but due to the detrimental effect on the design of the code (as each conditional block adds a certain complexity), we decided not to allow this for type 3R clones.

@@

\begin{equation}\label{eq:typerelation}
Cloned nodes T1R \subseteq Cloned nodes T2R \subseteq T3R
\end{equation}

### 2019-08-11 sections/results.tex
Proper capitalization for some concepts

\subsection{Relations Between Clone Instances} \label{chap:relationsinstances}
\todo{update these results}
In section \ref{sec:setuprelation} we introduced our experiments regarding relations between clones. Table~\ref{table:relations} contains our results regarding the relations between clone instances.

@@

\subsection{Clone instance location}\label{chap:clonelocation}
\todo{update these results}
Measuring the clone locations categories defined in section \ref{sec:setuplocation} yields the results displayed in Table \ref{table:contents}.

@@

\section{Clone instance contents}\label{chap:clonecontents}
\todo{update these results}
Measuring the clone contents categories defined in section \ref{sec:setupcontents} yields the results displayed in Table \ref{table:contents}.

@@

\section{Merging duplicate code through refactoring} \label{sec:refactorability}
\todo{update these results}
In section \ref{sec:refactorabilitysetup} we define categories regarding the extent to which clones can be refactored through method extraction. Our results are displayed in Table \ref{table:refactorability}.

>>>>

\subsection{Relations Between Clone Instances} \label{chap:relationsinstances}
\todo{update these results}
In Section~\ref{sec:setuprelation} we introduced our experiments regarding relations between clones. Table~\ref{table:relations} contains our results regarding the relations between clone instances.

@@

\subsection{Clone instance location}\label{chap:clonelocation}
\todo{update these results}
Measuring the clone locations categories defined in Section~\ref{sec:setuplocation} yields the results displayed in Table \ref{table:contents}.

@@

\section{Clone instance contents}\label{chap:clonecontents}
\todo{update these results}
Measuring the clone contents categories defined in Section~\ref{sec:setupcontents} yields the results displayed in Table \ref{table:contents}.

@@

\section{Merging duplicate code through refactoring} \label{sec:refactorability}
\todo{update these results}
In Section~\ref{sec:refactorabilitysetup} we define categories regarding the extent to which clones can be refactored through method extraction. Our results are displayed in Table \ref{table:refactorability}.
